Toward dynamic models of combat or lack thereof in animal mating

### 2016-03-18 ms.tex
logging of fights

@@

% whats b? define/ state value
Within the immature phase, we assume males grow with rate
\begin{equation} \label{eq:growth}
    \frac{dm}{dt} = a m^ b
\end{equation}

@@

% needs more detail, need location and width parameter values

We Pull the maturation time of a male randomly from the logistic distribution with location and width parameters.

% state k_im value and example probability
If we assume a constant mortality rate $k_{im}$ during the immature phase, the probability of a male surviving to time $t$ from $0$ is given by
\begin{equation} \label{eq:surv_immature}
    S(t) = \exp(-k_{im} t).
\end{equation}

When a male matures it is given an energy budget linearly proportional to its mass such that: 
$$ E = k_em$$
After a male matures it cannot gain more energy. When a mature male uses its energy allocation it dies.

@@

The energy used by a searching male with speed $v$ and radius $r$ is 
\begin{equation}
    E(r,v) = k_{search} \cdot r \cdot v
\end{equation}
where $k_{search}$ is a constant.  % state value

@@

The encounter rate $\lambda$ for an individual is:
\begin{equation} \label{eq:encounter rate}
    \lambda = \frac {N \cdot 2 \cdot r \cdot v} {A_{total}}
\end{equation}
\citep{Gurarie2012}
We assume spatially random nests so nest discovery is a poisson process with the the expected number of encounters in one time step being $\lambda$.
Because of this we can then use this as the rate parameter to repeatedly draw the time of the next event for this individual from the exponential distribution.
We do this for each male from when they mature up to when the females mature.
When an event occurs we select a nest randomly and then the male will either occupy the nest or contest another male for control of the nest.

@@

\subsection{Contests}
When a searching male encounters a nest that is occupied a contest over possession of that nest occurs. Contests occur instantaneously, taking up .

The costs of a contest escalates with more commitment.

We will assume an escalation curve $kappa$ such that:
$$\kappa : Commitment \rightarrow Intensity$$
where $Intensity$ is $\frac{Energy}{Commitment}$

We use the escalation curve $$ \kappa (c) = c^2 $$

At the start of the contest each male chooses how much  energy they are willing to commit.
They do this by reporting their level of ``commitment'' $c_i$.

The male that chooses to commit more of its energy to the contest will win with probability $= logit^{-1}(k(c_i - c_j))$. Where k is is a parameter that shapes how sensitive the contest is to small differences in commitment values.

The energy that this male chose to commit will then be deducted from both males with the loser continuing to search for mates

To get the cost of the fight we take the integral of the commitment function $\kappa$:

$$Energy Spent = \int_0^{c_i}{\kappa(t)}dt $$

if the $\kappa(c)$ is $c^2$ then
$$EnergySpent = \frac{c_i^3}{3}$$
Where $c_i$ is the losing male.

The male chooses how much energy to commit based off of a commitment function. for a male $i$ against a male $j$

An exponential one:
$$\mathrm{commitment} = e^\beta \cdot (\frac{m_i}{m_j})^\alpha $$

Where the trait values:
$$\alpha , \beta$$
Are under selection.

@@

\begin{enumerate}
    \item aggression 0 ($\beta$) called $e_0$
    \item aggression 1 ($\alpha$) called $k$

    \item r, the search radius of a searching male, See Fig. \ref{fig:area searched}.
    \item v, the search \"velocity" of a male.
\end{enumerate}

@@

\myfigure{figures/e_0_patch_area.png}{e 0 with patch area}{figures/clustered_e_0_plot.pdf}{an equivilant plot with k means clustering done, with the k parameter chosen by $pamk$, the number of clusters (coexisting strategies) can be be seen on the left}

>>>>

We use $F = 20$ and $N = 500$

@@

Within the immature phase, we assume males grow with rate
\begin{equation} \label{eq:growth}
    \frac{dm}{dt} = a m^ b
\end{equation}

@@

We use $a = 0.5$ and $b = 0.1$ and $m(0) = 5$

We Pull the maturation time of a male randomly from the logistic distribution with center $= 4$ and width $= 1$.

If we assume a constant mortality rate $k_{im}$ during the immature phase, the probability of a male surviving to time $t$ from $0$ is given by
\begin{equation} \label{eq:surv_immature}
    S(t) = \exp(-k_{im} t).
\end{equation}

We use $k_{im} = 0.2$

For example the probability of a male surviving to maturity from to time = $5$ is $exp(-0.2 \cdot 5)= 0.3678$.

When a male matures it is given an energy budget linearly proportional to its mass such that: 
$$ E = k_em$$

We use $k_e = 10$

After a male matures it cannot gain more energy. When a mature male uses its energy allocation it dies.

@@

The energy used by a searching male with speed $v$ and radius $r$ is 
\begin{equation}
    E(r,v) = k_{search} \cdot r \cdot v
\end{equation}
where $k_{search} = 5$ is a constant.  % state value

@@

The encounter rate $\lambda$ for an individual is:
\begin{equation} \label{eq:encounter rate}
    \lambda = \frac {N \cdot 2 \cdot r \cdot v} {A_{total}}
\end{equation}
\citep{Gurarie2012}

We assume spatially random nests so nest discovery is a poisson process with the the expected number of encounters in one time step being $\lambda$.
Because of this we can then use $\lambda$ as the rate parameter to repeatedly draw the time of the next event for this individual from the exponential distribution.

We do this for each male from when they mature up to when the females mature (at $\text{time} = 10$).
When an event occurs we select a nest randomly and then the male will either occupy the nest or contest another male for control of the nest.

@@

\subsection{Contests}
When a searching male encounters a nest that is occupied a contest over possession of that nest occurs.

Contests occur instantaneously, taking up no time, this means that another male cannot encounter the nest whilst the contest is taking place, additionally the contest has no intrinsic metabolic cost.

At the start of the contest each male chooses how much  energy they are willing to commit to this contest, they have full information about their opponents mass.

They do this by reporting their level of commitment as a positive value $c$.

The intensity of a fight increases with more commitment:

We define $\text{Intensity} = \frac{\text{d}Energy}{\text{d}Commitment}$

We use an intensity curve $I(c) = c^2$

\begin{figure}[h!]
    \centering
    \includegraphics[width=10cm,height=8cm]{figures/example_escalation.png}
    \caption{The energy spent in the fight between male i and male j is the pink shaded area.}
\end{figure}

To get the cost of the fight we take the integral of the intensity function:
$$E_\text{cost}(c) = \int_0^{c}{I(t)}dt = \int_0^{c}{t^2}dt $$

The lower energy cost ( corrosponding to $c_i$ in the diagram) will then be deducted from both males.

Male $i$ will loose the contest with probability 
$$P(\text{})= logit^{-1}(k(c_j - c_i))$$

Where k is is a parameter that shapes how sensitive the contest is to small differences in commitment values (we use $k = 10$).
can lower k be thought of as perception error (?? mike).

The male chooses how much energy to commit based off of a commitment function. For a male $i$ against a male $j$

$$\text{commitment} = e^\beta \cdot (\frac{m_i}{m_j})^\alpha $$

Where the trait values:
$$\alpha , \beta$$
are under selection.

@@

\begin{enumerate}
    \item $\beta$
    \item $\alpha$
    \item $r$, the search radius of a searching male, See Fig. \ref{fig:area searched}.
    \item $v$, the search speed of a male.
\end{enumerate}

The initial distribution of $\alpha$ and $\beta$ is $N(0, 3)$.

The initial distribution of $v$ and $r$ is $N(2, 0.5)$.

Where $N(\mu, \sigma)$, is the normal distribution with mean $= \mu$ and standard deviation $\sigma$.

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=15cm,height=10cm]{figures/clustered_e_0_plot.pdf}
    \caption{A plot of the distribution of the aggression parameter $beta$ over varing $A_{total}$ with k means clustering done, with the k parameter chosen by $pamk$, the number of clusters (coexisting strategies) can be be seen on the right}
\end{figure}
